Rubin Observatory Pointing Algorithm and Simulations

### 2025-01-30 ms.tex
Updates from Overleaf

\title{Rubin Observatory Scheduling Algorithm and Simulations}

@@

\author[0000-0003-2874-6464]{Peter Yoachim}
\affiliation{Department of Astronomy and the DiRAC Institute, University of Washington, Seattle, WA, USA}

\author[0000-0001-5916-0031]{R. Lynne Jones}
\affiliation{Department of Astronomy and the DiRAC Institute, University of Washington, Seattle, WA 98195, USA}

\author[0000-0002-7357-0317]{Eric H. Neilsen}
\affiliation{Fermi National Accelerator Laboratory, P.O. Box 500, Batavia, IL 60510, USA)}

\author{Tiago Ribeiro}
\affiliation{Vera C. Rubin Observatory (United States)}

@@

@@

@@

including scheduled and unscheduled downtime

@@

For clouds, we assume that observations are not possible if more than 30\% of the sky is cloudy. This level is reached 28\% of the time. XXX-mention how this compares to other weather downtimes. Ugh, now I can't find that nice SOAR page that had a nice table.

@@

Run through the information the model observatory generates and passes to the scheduler.

There is an additional scheduling algorithm that controls which filters are loaded at the start of each night. This is configured so that if the moon is 40\% illuminated or more the $grizy$ filters are loaded, otherwise the $ugriy$ filters are loaded.

@@

% Might be enough to discuss here to break into more figures
\begin{figure}
    \epsscale{0.5}
    \plotone{plots/cond_airmass.pdf}\plotone{plots/cond_seeing.pdf} \\
    \plotone{plots/cond_selong.pdf}\plotone{plots/cond_skybright.pdf} \\
    \plotone{plots/cond_slewtime.pdf}
    \epsscale{1}
    \caption{Information generated by the model observatory and passed to the scheduling algorithm. These are Mollweide projections of RA,dec coordinates for October 1, 2023. Areas in gray are unreachable by the telescope because they are low altitude or in the zenith exclusion zone. The sun is setting with an altitude of -17 degrees. All the maps are stored as nside=32 HEALpix arrays. }
    \label{fig:conditions}
\end{figure}

\section{Deep Drilling Fields}

The LSST is planned to include several deep drilling fields. xxx--each DDF, maybe a cite for other studies of them. We plan to observe the Euclid Deep Field South (EDFS) as a pair of pointings.

@@

\begin{table}
    \centering
        \begin{tabular}{lcc}
        \toprule
        Short Name &      RA (deg) &     Dec (deg) \\
        \hline
           ELAISS1 &   9.450 & -44.000 \\
           XMM\_LSS &  35.708 &  -4.750 \\
             ECDFS &  53.125 & -28.100 \\
            COSMOS & 150.100 &   2.182 \\
            EDFS\_a &  58.900 & -49.315 \\
            EDFS\_b &  63.600 & -47.600 \\
        \hline
        \end{tabular}
    \caption{Adopted locations for Deep Drilling Fields.}
    \label{tab:ddf_loc}
\end{table}

@@

\section{Markov Decision Process}

@@

xxx-maybe make a plot of the baseline tesselation that gets randomized.

@@

\software{Astropy \citep{astropy2013, astropy2018}, 
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array},
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022},
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009},
          GNU Parallel \citep{Tange2023}}

@@

\begin{acknowledgments}
This work was facilitated through the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system at the University of Washington
\end{acknowledgments}

>>>>

\title{Rubin Observatory Pointing Algorithm and Simulations}

@@

\author[0000-0003-2874-6464]{Peter Yoachim}
\affiliation{Department of Astronomy and the DiRAC Institute, University of Washington, Seattle, WA 98195, USA}

\author[0000-0001-5916-0031]{R. Lynne Jones}
\affiliation{Rubin Observatory, 950 N. Cherry Ave., Tucson, AZ 85719, USA}
\affiliation{Aerotek, Suite 150, 4321 Still Creek Dr., Burnaby, BC V5C6S, Canada}

\author[0000-0002-7357-0317]{Eric H. Neilsen}
\affiliation{Fermi National Accelerator Laboratory, P.O. Box 500, Batavia, IL 60510, USA}

\author{Tiago Ribeiro}
\affiliation{Rubin Observatory, 950 N. Cherry Ave., Tucson, AZ 85719, USA}

@@

xxx-Cite Eric's paper and LCOGT on integer programming.

Lots of buzz around AI models, but they are overblown \citep{Hicks24}

@@

the overview paper \citet{Rubin_overview2019}

@@

Optional to include a finite jerk for each component using motion profiles from \citet{Mu2009}.

including scheduled and unscheduled downtime---xxx details of each.

@@

For clouds, we assume that observations are not possible if more than 30\% of the sky is cloudy. This level is reached 26\% of the time. XXX-mention how this compares to other weather downtimes. Ugh, now I can't find that nice SOAR page that had a nice table.

@@

Run through the information the model observatory generates and passes to the decision maker.

There is an additional algorithm that controls which filters are loaded at the start of each night. This is configured so that if the moon is 40\% illuminated or more the $grizy$ filters are loaded, otherwise the $ugriy$ filters are loaded.

@@

% Might be enough to discuss here to break into more figures
\begin{figure}
    \epsscale{0.5}
    \plotone{plots/cond_airmass.pdf}\plotone{plots/cond_seeing.pdf} \\
    \plotone{plots/cond_selong.pdf}\plotone{plots/cond_skybright.pdf} \\
    \plotone{plots/cond_slewtime.pdf}
    \epsscale{1}
    \caption{Information generated by the model observatory and passed to the observing decision algorithm. These are Mollweide projections of RA,dec coordinates for October 1, 2023. Areas in gray are unreachable by the telescope because they are low altitude or in the zenith exclusion zone. The sun is setting with an altitude of -17 degrees. All the maps are stored as nside=32 HEALpix arrays. }
    \label{fig:conditions}
\end{figure}

\section{Survey Modes}

\subsection{Deep Drilling Fields}
The LSST is planned to include several deep drilling fields. xxx--each DDF, maybe a cite for other studies of them. We plan to observe the Euclid Deep Field South (EDFS) as a pair of pointings.

@@

% Generated in DDF_table.ipynb
\begin{table}
    \centering
        \begin{tabular}{lrr}
            \toprule
            Short Name & RA & Dec \\
            \midrule
            ELAISS1 & 9.450 & -44.025 \\
            XMM LSS & 35.575 & -4.817 \\
            ECDFS & 52.979 & -28.117 \\
            COSMOS & 150.108 & 2.234 \\
            EDFS a & 58.900 & -49.320 \\
            EDFS b & 63.600 & -47.600 \\
            \bottomrule
            \end{tabular}
    \caption{Adopted locations for Deep Drilling Fields.}
    \label{tab:ddf_loc}
\end{table}

@@

\subsection{Markov Decision Process}

@@

\begin{figure}
    \plotone{plots/hp2p_n1.pdf}
    \plotone{plots/hp2p_n5.pdf}
    \plotone{plots/nover_n1.pdf}
    \plotone{plots/nover_n5.pdf}
\end{figure}

@@

Twilight survey for inner solar system objects, similar to \citet{Ye2020}.

\section{Conclusions}

\software{Astropy \citep{astropy2013, astropy2018}, 
          Healpy and HEALpix\footnote{\url{http://healpix.sourceforge.net}} \citep{healpix2005, Zonca2019},
          Numpy \citep{harris2020array},
          Matplotlib \citep{Hunter:2007},
          Scipy \citep{2020SciPy-NMeth},
          Pandas \citep{reback2022},
          Scikit-learn \citep{scikit-learn},
          OpenOrb \citep{Granvik2009},
          GNU Parallel \citep{Tange2023},
          rubin\_scheduler \citep{rubin_scheduler2024},
          rubin\_sim \citep{rubin_sim2024}}

@@

\begin{acknowledgments}
This work was facilitated through the use of advanced computational, storage, and networking infrastructure provided by the Hyak supercomputer system at the University of Washington.
\end{acknowledgments}
